---
title: "Data Visualization"
description: |
  Companion website for the MSP Data Visualization course offered in Fall 2023 Mini 1 and taught by Professor Ron Yurko.
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

# Learn more about publishing to GitHub Pages at:
# https://rstudio.github.io/distill/publish_website.html#github-pages

```


## Lectures

Lecture slides are available below in html format, and the .Rmd files are linked from GitHub.

| Lecture | Topic                         | html                                                                               | Rmd                                                                                                              |
|-----------|-----------------------------------------|-----------|-----------|
| 1       | Introduction and the Grammar of Graphics | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/slides/Lec1#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/slides/Lec1.Rmd) |
| 2      | 1D Categorical Data | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/slides/Lec2#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/slides/Lec2.Rmd) |
| 3      | 2D Categorical Data | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/slides/Lec3#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/slides/Lec3.Rmd) |
| 4      | 1D Quantitative Data | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/slides/Lec4#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/slides/Lec4.Rmd) |
| 5      | Density Estimation | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/slides/Lec5#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/slides/Lec5.Rmd) |
| 6      | 2D Quantitative Data | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/slides/Lec6#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/slides/Lec6.Rmd) |
| 7      | More 2D Quant. and Intro to High Dim. Data | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/slides/Lec7#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/slides/Lec7.Rmd) |
| 8      | High Dimensional Data | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/slides/Lec8#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/slides/Lec8.Rmd) |
| 9      | More High Dim. Data | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/slides/Lec9#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/slides/Lec9.Rmd) |
| 10      | Shiny and Visualizing Text | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/slides/Lec10#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/slides/Lec10.Rmd) |
| 11      | Visualizing Spatial Data and Maps | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/slides/Lec11#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/slides/Lec11.Rmd) |
| 12      | Visualizing Trends and Animations | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/slides/Lec12#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/slides/Lec12.Rmd) |



## R Demos

Additional `R` demos are available below in html format.

| Demo | Topic                         | html                                                                               | Rmd                                                                                                              |
|-----------|-----------------------------------------|-----------|-----------|
| 0       | Installing R and RStudio | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/demos/Demo0.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/demos/Demo0.Rmd) |
| 1       | Into the tidyverse | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/demos/Demo1.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/demos/Demo1.Rmd) |
| 2       | Scatterplots and regression	 | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/demos/Demo2.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/demos/Demo2.Rmd) |
| 3       | Pairs plots, correlograms, and parallel coordinates		 | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/demos/Demo3.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/demos/Demo3.Rmd) |
| 4       | Multi-Dimensional Scaling	 | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/demos/Demo4.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/demos/Demo4.Rmd) |
| 5       | Principal Component Analysis	 | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/demos/Demo5.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/demos/Demo5.Rmd) |
| 6       | Intro to t-SNE and UMAP	 | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/demos/Demo6.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/demos/Demo6.Rmd) |
| 7      | Visualizing Text Data		 | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/demos/Demo7.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/demos/Demo7.Rmd) |
| 8      | Visualizing Spatial Point Data	 | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/demos/Demo8.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/demos/Demo8.Rmd) |
| 9      | Visualizing Areal Data | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/demos/Demo9.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/demos/Demo9.Rmd) |
| 10     | Visualizing Trends | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/demos/Demo10.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/demos/Demo10.Rmd) |
| 11     | Animations | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/demos/Demo11.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/demos/Demo11.Rmd) |
| 12     | Modifying Colors and Themes | [`r icons::fontawesome$solid$laptop`](https://cmu-36613-fall23.netlify.app/demos/Demo12.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/DataViz-36613-Fall23/blob/main/demos/Demo12.Rmd) |





## Resources

### Recommended textbooks

-   R for Data Science: <https://r4ds.hadley.nz/>

-   Fundamentals of Data Visualization: <https://clauswilke.com/dataviz/index.html>

-   Data Visualization: A Practical Introduction: <https://socviz.co/index.html#preface>

-   ggplot2: Elegant Graphics for Data Analysis: <https://ggplot2-book.org/>

-   Mastering Shiny: <https://mastering-shiny.org/index.html>

-   The best resource for working with `git` and GitHub in `R` / RStudio: <https://happygitwithr.com/>

### More `ggplot2` resources

-   The `ggplot2` cheatsheet: <https://www.maths.usyd.edu.au/u/UG/SM/STAT3022/r/current/Misc/data-visualization-2.1.pdf>

-   Creating custom themes: <https://themockup.blog/posts/2020-12-26-creating-and-using-custom-ggplot2-themes/>

-   R Graphics cookbook: <https://r-graphics.org/>

-   A guide to various scales: <https://ggplot2tor.com/scales>

### R Markdown

-   The definitive guide to R Markdown: <https://bookdown.org/yihui/rmarkdown/>

-   R Markdown reference guide: <https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf>

-   R Markdown cheat sheet: <https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf>

## Acknowledgments {.appendix}

This course website is made with the [distill package](https://rstudio.github.io/distill/). Thanks to [Meghan Hall](https://meghan.rbind.io/), [Alison Hill](https://rstudio-education.github.io/teaching-in-production/), and [Shamindra Shrotriya](https://www.shamindras.com/posts/2019-07-11-shrotriya2019distillpt1/) for their incredible examples and tutorials that helped in the process of making this site.

```{r, include=FALSE, results='asis'}
library(metathis)
meta() %>%
  meta_social(
    title = "CMU 36-613",
    description = "Course website for Data Visualization, by Ron Yurko",
    url = "https://cmu-36613.netlify.app/",
    image_alt = "CMU 36-613",
    og_type = "website",
    og_author = "Ron Yurko",
    twitter_card_type = "summary",
    twitter_creator = "@Stat_Ron"
  )
```
