---
title: "36-613: Data Visualization"
subtitle: "Extras with htmlwidgets and Dashboards"  
author: 
  - "Professor Ron Yurko"
date: '10/12/2022'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE,
  fig.path = "figs/Lec11/"
)
xaringanExtra::use_scribble()
xaringanExtra::use_clipboard()
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#c41230",
  secondary_color = "#4d5051",
  inverse_header_color = "#FFFFFF"
)
```

```{r, echo = FALSE}
library(tidyverse)
library(palmerpenguins)
data(penguins)
```


# REMINDER: Please fill out the FCE!

#### Faculty course evaluations (FCEs) are out now

#### I take feedback very seriously and I want the course to be useful!

#### If you enjoyed this class, please fill out the FCE.

#### If you didn’t enjoy this class, please fill out the FCE.


---

## A bridge between `R` and JavaScript

-  JavaScript enables web developers to create __client-side__ web applications

  - Computations are happening __in the client's browser__, instead of the host's web servers
  
- `D3` (or `D3.js`) is the most popular JavaScript library for client-side dynamic data visualizations

  - `D3` == 'data-driven documents'

--

- RStudio developers created the [`htmlwidgets`](http://www.htmlwidgets.org/index.html) enabling `R` users to use `D3` without needing to learn JavaScript!

- Renders data visualizations in HTML using `D3` wrappers

  - [Incredible gallery with a variety of uses](http://gallery.htmlwidgets.org/), including a chess package called [`rchess`](http://jkunst.com/rchess/)!
  
---

## [Leaflet](https://rstudio.github.io/leaflet/): interactive HTML maps

.pull-left[

```{r cmu-map, eval = FALSE}
cmu <- 
  tibble(address = "Carnegie Mellon University, Pittsburgh, PA") %>%
  tidygeocoder::geocode(address, method = "osm") #<<

library(leaflet)
leaflet() %>% addTiles() %>% #<<
  addMarkers(data = cmu) #<<
```

- `addTiles()`: builds layer with static map (default from OpenStreetMap)

- `addMarkers()`: add marker at point location, e.g., CMU

- Use [`tidygeocoder`](https://cran.r-project.org/web/packages/tidygeocoder/vignettes/tidygeocoder.html) for spatial queries


]

.pull-right[

```{r ref.label="cmu-map", fig.align='center', fig.height=7, echo = FALSE}

```



]

---

## [DataTables (DT) package](https://rstudio.github.io/DT/) for interactive tables

```{r, fig.align='center', out.width="80%"}
library(DT)
datatable(penguins[, 1:6], options = list(pageLength = 5)) #<<
```



---

## Interactive plots with [Plotly](https://plotly.com/r/)

.pull-left[

- Start with an initial plain plot

```{r plain-plot, eval = FALSE}
scatter_plain <- penguins %>% 
  ggplot(aes(x = body_mass_g, y = bill_length_mm, 
             color = species)) +
  geom_point(alpha = 0.5, size = 2) +
  labs(x = "Body Mass (g)", y = "Bill Length (mm)") +
  theme_bw()
scatter_plain
```


]
.pull-right[

```{r ref.label="plain-plot", fig.align='center', fig.height=7, echo = FALSE}

```


]

---

## Interactive plots with [Plotly](https://plotly.com/r/)

.pull-left[

- Make it interactive with `ggplotly`

```{r go-plotly, eval = FALSE}
library(plotly)
ggplotly(scatter_plain)
```

- `plotly` has its own plotting syntax, but we can just use `ggplot2` instead 

  - there will be some annoying differences at times, so just be careful
  
- Several ways to interact: filter, zoom, and view additional info by hovering over the points with tooltip...

]
.pull-right[

```{r ref.label="go-plotly", fig.align='center', fig.height=7, echo = FALSE}

```
]

---

## Customize the tooltip

.pull-left[

- Update to include `text` with penguin's `sex`

```{r tool-plotly, eval = FALSE}
scatter_upd <- penguins %>% 
  ggplot(aes(x = body_mass_g, y = bill_length_mm, 
             color = species,
             text = paste("sex:", sex))) + #<<
  geom_point(alpha = 0.5, size = 2) +
  labs(x = "Body Mass (g)", y = "Bill Length (mm)") +
  theme_bw()

ggplotly(scatter_upd, 
         tooltip = c("text", "species")) #<<
```

- We can specify what is included in the tooltip when hovering over the points

]
.pull-right[

```{r ref.label="tool-plotly", fig.align='center', fig.height=7, echo = FALSE}

```

]

---

# Putting it all together with dashboards

- Dashboards are popular way to make data and visualizations available to clients, managers, stakeholders, etc. to help with decision making

- Typically include a mix of graphics and text, depending on the context 

--

- Can easily make HTML dashboards using the [`flexdashboard`](https://pkgs.rstudio.com/flexdashboard/) package

  - Provide interactive capabilities within an HTML file (i.e., you can email this!)
  
  - __These are NOT fully interactive like Shiny apps__
  
  - [You can use Shiny within `flexdashboard`](https://pkgs.rstudio.com/flexdashboard/articles/shiny.html) but it requires a Shiny server
  
  - Another option is [`shinydashboard`](https://rstudio.github.io/shinydashboard/)
  
- Outside of `R`, [Tableau](https://public.tableau.com/app/discover) is a popular tool (can use free Tableau Public version)


---
class: center, middle

# DO IT LIVE

---
class: center, middle

# Final Report due Friday Oct 14th by 5 PM EDT via email! (just one per group)

Recommended reading: 

[Modern Data Science with R: CH 14 Dynamic and customized data graphics](https://mdsr-book.github.io/mdsr2e/ch-vizIII.html)

[`plotly` resources](https://plotly.com/r/)

[`htmlwidgets` gallery](http://gallery.htmlwidgets.org/)




