---
title: "36-613: Data Visualization"
subtitle: "More 2D Quant. and Intro to High Dimensional Data"  
author: 
  - "Professor Ron Yurko"
date: '9/20/2023'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE,
  fig.path = "figs/Lec7/"
)
xaringanExtra::use_clipboard()
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#c41230",
  secondary_color = "#0277BD",
  inverse_header_color = "#FFFFFF"
)
```


```{r, include = FALSE}
library(tidyverse)
library(palmerpenguins)
```



## Visuals to focus on the joint distribution

```{r, echo = FALSE}
lebron_shots <- read_csv("https://raw.githubusercontent.com/ryurko/Spring23-36315-Data/main/lebron_shots.csv")
```

.pull-left[

- Example [dataset of LeBron James' shots this NBA season](https://raw.githubusercontent.com/ryurko/Spring23-36315-Data/main/lebron_shots.csv) accessed using [`hoopR`](https://hoopr.sportsdataverse.org/)

```{r lebron-shots, eval = FALSE}
lebron_shots %>%
  ggplot(aes(x = coordinate_x, 
             y = coordinate_y)) +
  geom_point(alpha = 0.4) +
  coord_fixed() + #<<
  theme_bw()
```

- Where are the high/low concentrations of X,Y?

- How do we display concentration for 2D data?

- `coord_fixed()` so axes match with unit scales

]

.pull-right[
```{r ref.label="lebron-shots", echo = FALSE, fig.height=8}
```

]

---

## Display 2D contour plot

.pull-left[


```{r lebron-contour, eval = FALSE}
lebron_shots %>%
  ggplot(aes(x = coordinate_x, 
             y = coordinate_y)) +
  geom_point(alpha = 0.4) +
  geom_density2d(binwidth = 0.0001) + #<<
  coord_fixed() +
  theme_bw()
```

- Use `geom_density2d` to display contour lines

- Inner lines denote "peaks"

]

.pull-right[
```{r ref.label="lebron-contour", echo = FALSE, fig.height=8}
```

]





---

## Back to the penguins...

Pretend I give you this `penguins` dataset and I ask you to make a plot __for every pairwise comparison__...

```{r}
library(palmerpenguins)
penguins %>% slice(1:3)
```

--

We can create a __pairs plot__ to see __all__ pairwise relationships __in one plot__

Pairs plot can include the various kinds of pairwise plots we've seen:

- Two quantitative variables: scatterplot

- One categorical, one quantitative: side-by-side violins, stacked
histograms, overlaid densities

- Two categorical: stacked bars, side-by-side bars, mosaic plots

---

## Pairs plots for penguins

.pull-left[

Use the [`GGally`](https://ggobi.github.io/ggally/index.html) package

```{r pairs-plot, eval = FALSE}
library(GGally)
penguins %>% 
  ggpairs(columns = 3:6) #<<
```

Main arguments to change are:

+ `data`: specifies the dataset

+ `columns`: columns of data you want in the plot (can specify with vector of column names or numbers referring to the column indices)

+ `mapping`: aesthetics using `aes()` - most important is `aes(color = <variable name>)`

Created pairs plot above by specifying `columns` as the four columns of continuous variables (columns 3-6)


]

.pull-right[
```{r ref.label="pairs-plot", echo = FALSE, fig.height=7}

```


]

---

## Pairs plots for penguins

.pull-left[

Annoying aspect: change `alpha` directly with `aes` when using `ggpairs`:

```{r pairs-plot-alpha, eval = FALSE}
penguins %>% 
  ggpairs(columns = 3:6,
          mapping = aes(alpha = 0.5)) #<<
```

- Diagonal: marginal distributions

- Off-diagonal: joint (pairwise) distributions or statistical summaries (e.g., correlation)

- Matrix of plots is symmetric

]

.pull-right[
```{r ref.label="pairs-plot-alpha", echo = FALSE, fig.height=7}

```


]

---

## Read Demo3 for more info on customization!

.pull-left[

```{r, echo = FALSE, fig.height=7}
penguins %>% 
  ggpairs(columns = c("bill_length_mm", "body_mass_g", "island"),
          mapping = aes(alpha = 0.5, color = species), 
          lower = list(
            continuous = "smooth_lm", 
            combo = "facetdensitystrip"
          ),
          upper = list(
            continuous = "cor",
            combo = "facethist"
          )
  )
```


]

.pull-right[

```{r, echo = FALSE, fig.height=7}
penguins %>%
  dplyr::select(species, body_mass_g, ends_with("_mm")) %>%
  ggpairs(mapping = aes(color = species, alpha = 0.5),
          columns = c("flipper_length_mm", "body_mass_g",
                      "bill_length_mm", "bill_depth_mm")) +
  scale_colour_manual(values = c("darkorange","purple","cyan4")) +
  scale_fill_manual(values = c("darkorange","purple","cyan4")) +
  theme_bw() +
  theme(strip.text = element_text(size = 7))
```

]


---

## What about high-dimensional data?

Consider this [dataset]((https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-12-21/readme.md) containing nutritional information about Starbucks drinks:

```{r, warning = FALSE, message = FALSE}
starbucks <- 
  read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-12-21/starbucks.csv") %>%
  # Convert columns to numeric that were saved as character
  mutate(trans_fat_g = as.numeric(trans_fat_g), fiber_g = as.numeric(fiber_g))
starbucks %>% slice(1)
```


#### How do we visualize this dataset? 

--

- Tedious task: make a series of pairs plots (one giant pairs plot would overwhelming)


---

## What about high-dimensional data?


```{r, warning = FALSE, message = FALSE}
starbucks <- 
  read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-12-21/starbucks.csv") %>%
  # Convert columns to numeric that were saved as character
  mutate(trans_fat_g = as.numeric(trans_fat_g), fiber_g = as.numeric(fiber_g))
starbucks %>% slice(1)
```

#### Goals to keep in mind with visualizing high-dimensional data

- __Visualize structure among observations__ using distances matrices, projections (Monday's lecture)

- __Visualize structure among variables__ using correlation as "distance"


---

## Correlogram to visualize correlation matrix

.pull-left[

Use the [`ggcorrplot`](https://rpkgs.datanovia.com/ggcorrplot/) package

```{r cor-plot, eval = FALSE}
starbucks_quant_cor <- 
  cor(dplyr::select(starbucks, #<<
                    serv_size_m_l:caffeine_mg))

library(ggcorrplot)
ggcorrplot(starbucks_quant_cor, #<<
           method = "circle",
           hc.order = TRUE,
           type = "lower")
```

- Compute the correlation matrix (using quantitative variables)

- Can rearrange using `hc.order = TRUE` based on clustering

- See Demo3 for more examples...

]

.pull-right[

```{r ref.label="cor-plot", echo = FALSE, fig.height=7}

```


]

---

## Parallel coordinates plot with [`ggparcoord`](https://ggobi.github.io/ggally/reference/ggparcoord.html)

.pull-left[

- Display each variable side-by-side on standardized axis

- Connect observations with lines

```{r par-coord, eval = FALSE}
starbucks %>%
  ggparcoord(columns = 5:15, #<<
             alphaLines = .1) + #<<
  theme(axis.text.x = 
          element_text(angle = 90))
```

- Can change `scale` method for y-axis

- Useful for moderate number of observations and variables

- __How do we order the x-axis?__

- __Does this agree with the correlogram?__

]
.pull-right[

```{r ref.label="par-coord", echo = FALSE, fig.height=7}

```

]

---
class: center, middle

# Next time: More High-Dimensional Data

Reminder: __HW3 due tonight!__ 

Recommended reading: 

[CW Chapter 12 Visualizing associations among two or more quantitative variables](https://clauswilke.com/dataviz/visualizing-associations.html)


