---
title: "Modifying Colors and Themes"
author: "Professor Ron Yurko"
output: html_document
---


###  The graphs below don't have proper titles, axis labels, legends, etc.  Please take care to do this on your own graphs.

Throughout this demo we will once again use the [`palmerpenguins` dataset](https://allisonhorst.github.io/palmerpenguins/articles/intro.html). To access the data, you will need to install the `palmerpenguins` package:

```{r, eval = FALSE}
install.packages("palmerpenguins")
```

```{r}
library(tidyverse)
library(palmerpenguins)
data(penguins)
```


## Notes on colors in plots

Three types of color scales to work with:

1. __Qualitative__: distinguishing discrete items that don't have an order (nominal categorical). Colors should be distinct and equal with none standing out unless otherwise desired for emphasis.

+ Do __NOT__ use a discrete scale on a continuous variable

2. __Sequential__:  when data values are mapped to one shade, e.g., in a choropleth, for an ordered categorical variable or low to high continuous variable

+ Do __NOT__ use a sequential scale on an unordered variable

3. __Divergent__: think of it as two sequential scales with a natural midpoint
midpoint could represent 0 (assuming +/- values) or 50% if your data spans the full scale

+ Do __NOT__ use a divergent scale on data without natural midpoint


### Options for `ggplot2` colors

The default color scheme is pretty bad to put it bluntly, but `ggplot2` has ColorBrewer built in which makes it easy to customize your color scales. For instance, we change the palette for the species plot from before.

```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_point(alpha = 0.5, size = 2) +
  scale_color_brewer(palette = "Set2") +
  labs(x = "Body Mass (g)", y = "Bill Length (mm)") +
  theme_bw()
```

Something you should keep in mind is to pick a [color-blind friendly palette](http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/). One simple way to do this is by using the `ggthemes` package which has color-blind friendly palettes included:

```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_point(alpha = 0.5, size = 2) +
  ggthemes::scale_color_colorblind() +
  labs(x = "Body Mass (g)", y = "Bill Length (mm)") +
  theme_bw()
```

In terms of displaying color from low to high, the [viridis scales](https://ggplot2.tidyverse.org/reference/scale_viridis.html) are excellent choices (and are also color-blind friendly!).

```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g, y = bill_length_mm, 
             color = flipper_length_mm)) +
  geom_point(alpha = 0.5, size = 2) +
  scale_color_viridis_c() +
  labs(x = "Body Mass (g)", y = "Bill Length (mm)",
       color = "Flipper Length (mm)") +
  theme_bw()
```


## Notes on themes

We have not explicitly talked about this throughout the semester, but you have seen various changes to the `theme` of plots for customization. __You will constantly be changing the theme of your plots to optimize the display.__ Fortunately, there are a number of built-in themes you can use to start with rather than the default `theme_gray()`:

```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_point(alpha = 0.5, size = 2) +
  ggthemes::scale_color_colorblind() +
  labs(x = "Body Mass (g)", y = "Bill Length (mm)") +
  theme_gray()
```

For instance, you have seen me use `theme_bw()` many times throughout the semester:

```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_point(alpha = 0.5, size = 2) +
  ggthemes::scale_color_colorblind() +
  labs(x = "Body Mass (g)", y = "Bill Length (mm)") +
  theme_bw()
```

There are options such as `theme_minimal()`:

```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_point(alpha = 0.5, size = 2) +
  ggthemes::scale_color_colorblind() +
  labs(x = "Body Mass (g)", y = "Bill Length (mm)") +
  theme_minimal()
```


or `theme_classic()`:

```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_point(alpha = 0.5, size = 2) +
  ggthemes::scale_color_colorblind() +
  labs(x = "Body Mass (g)", y = "Bill Length (mm)") +
  theme_classic()
```


There are also packages with popular, such as the `ggthemes` package which includes, for example, `theme_economist()`:

```{r}
library(ggthemes)
penguins %>% 
  ggplot(aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_point(alpha = 0.5, size = 2) +
  ggthemes::scale_color_colorblind() +
  labs(x = "Body Mass (g)", y = "Bill Length (mm)") +
  theme_economist()
```

and `theme_fivethirtyeight()` to name a couple:

```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_point(alpha = 0.5, size = 2) +
  ggthemes::scale_color_colorblind() +
  labs(x = "Body Mass (g)", y = "Bill Length (mm)") +
  theme_fivethirtyeight()
```

With any theme you have picked, you can then modify specific components directly using the `theme()` layer. There are [many aspects of the plot's theme to modify](https://ggplot2.tidyverse.org/reference/theme.html), such as my decision to move the legend to the bottom of the figure, drop the legend title, and increase the font size for the y-axis:

```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_point(alpha = 0.5, size = 2) +
  ggthemes::scale_color_colorblind() +
  labs(x = "Body Mass (g)", y = "Bill Length (mm)",
       title = "Larger penguins tend to have larger bills",
       subtitle = "Positive relationship between mass and length is consistent across species") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        axis.text.y = element_text(size = 14),
        axis.text.x = element_text(size = 6))
```


If you're tired of explicitly customizing every plot in the same way all the time, then you should make a custom theme. It's quite easy to make a custom theme for `ggplot2` and of course [there are an incredible number of ways to customize your theme](https://themockup.blog/posts/2020-12-26-creating-and-using-custom-ggplot2-themes/). In the code chunk, I modify the `theme_bw()` theme using the `%+replace%` argument to make my new theme named `my_theme()` - which is stored as a function:

```{r}
my_theme <- function () {
  # Start with the base font size
  theme_bw(base_size = 10) %+replace%
    theme(
      panel.background  = element_blank(),
      plot.background = element_rect(fill = "transparent", color = NA), 
      legend.position = "bottom",
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.key = element_rect(fill = "transparent", color = NA),
      axis.ticks = element_blank(),
      panel.grid.major = element_line(color = "grey90", size = 0.3), 
      panel.grid.minor = element_blank(),
      plot.title = element_text(size = 18, 
                                hjust = 0, vjust = 0.5, 
                                face = "bold", 
                                margin = margin(b = 0.2, unit = "cm")),
      plot.subtitle = element_text(size = 12, hjust = 0, 
                                   vjust = 0.5, 
                                   margin = margin(b = 0.2, unit = "cm")),
      plot.caption = element_text(size = 7, hjust = 1,
                                  face = "italic", 
                                  margin = margin(t = 0.1, unit = "cm")),
      axis.text.x = element_text(size = 13),
      axis.text.y = element_text(size = 13)
    )
}
```


Now I can go ahead and my plot from before with this theme:

```{r}
penguins %>% 
  ggplot(aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_point(alpha = 0.5, size = 2) +
  ggthemes::scale_color_colorblind() +
  labs(x = "Body Mass (g)", y = "Bill Length (mm)",
       title = "Larger penguins tend to have larger bills",
       subtitle = "Positive relationship between mass and length is consistent across species") +
  my_theme()
```


